---
title: "Exercises HA 2.1 and 2.3"
author: "Alain Kuiete"
date: "6/4/2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 2.1 
Use the help function to explore what the series gold, woolyrnq and gas represent.

Use autoplot() to plot each of these in separate plots.
What is the frequency of each series? Hint: apply the frequency() function.
Use which.max() to spot the outlier in the gold series. Which observation was it?

Load the libraries
```{r}
library(fpp2)
```


```{r}
help(gold)
```
the forcast gold is a time series data that gives the daily morning prices from January 1st 1985 to March 31st 1989.

```{r}
tsdisplay(gold)
```

The autucorrelation plot shows that the data is  trended, cyclic, but the seasonality is not accentuated to be considered. It is part of white noise.
```{r}
autoplot(gold)
```



```{r}
frequency(gold)
```
 This confirms that the time series is daily
Outlier
```{r}
which.max(gold)
```
The maximun price of gold was on February 10 1987


```{r}
gold[770]
```
Since the gold frequency is 1, the outlier occure on February, 10 1987 with the morning daily price of $593.7




```{r}
help(woolyrnq)
```
woolyrnq is the quaterly production of woollen yarn in Australia.
The unit of production is in tonnes. The time series data is from Mars 1965 to September 1994.

```{r}
tsdisplay(woolyrnq)
```

```{r}
autoplot(woolyrnq)
```
The quarterly production of woollen shows cyclic

```{r}
frequency(woolyrnq)
```
This time series is quaterly

Outlier
```{r}
which.max(woolyrnq)
```
This maximun occur at the second quarter of 1969


```{r}

```


```{r}
which.min(woolyrnq)
```
The minimun is at the second  quarter of 1975

```{r}
help(gas)
```
The gas time series is the Australian monthly gas production from 1956 to 1995.


```{r}
tsdisplay(gas)
```

```{r}
autoplot(gas)
```


```{r}
frequency(gas)
```



```{r}
which.max(gas)
```



```{r}
which.max(gas)/frequency(gas) + 1956
```



```{r}
ggseasonplot(gas)
```

The seasonality of gas is changing over time progressively displaying a maximun on July or August


### Exercise 2.3
Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:


retaildata <- readxl::read_excel("retail.xlsx", skip=1)
The second argument (skip=1) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

myts <- ts(retaildata[,"A3349873A"],
  frequency=12, start=c(1982,4))
Explore your chosen retail time series using the following functions:

autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?


Load the dataset
```{r}
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
```

Turnover ;  New South Wales ;  Electrical and electronic goods retailing ;
```{r}
myts <- ts(retaildata[,"A3349336V"],
  frequency=12, start=c(1982,4))
```

```{r}
str(myts)
```


autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf()
```{r}
autoplot(myts)
```



```{r}
ggseasonplot(myts, polar = TRUE)
```
The polar coordinate shows that there is a large jump on sales in December. The month of June presents also a smooth increase in sales. Conversely, there are fall on sales in Februry and August.

```{r}
ggsubseriesplot(myts)
```


```{r}
gglagplot(myts)
```

```{r}
ggAcf(myts)
```

The lag 1, 12, and 24 are correlated. Signe of seasonality.























































