---
title: "data_624_hw_1"
author: "Jimmy Ng"
date: "6/15/2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages <- c("tidyverse", "fpp2", "forecast", "kableExtra", "broom")
pacman::p_load(char = packages)
```

## Hyndman 7.1 {.tabset .tabset-fade .tabset-pills}

Consider the `pigs` series — the number of pigs slaughtered in Victoria each month.

a. Use the `ses()` function in R to find the optimal values of $\alpha$ and $\ell_0$, and generate forecasts for the next four months.

b. Compute a 95% prediction interval for the first forecast using $\hat{y}\pm1.96s$ where $s$ is the standard deviation of the residuals. Compare your interval with the interval produced by R.

### a. ses() function
```{r echo = TRUE, collapse = TRUE, fig.width = 8, fig.height = 10}
# take a look of the ts object
pigs <- read.csv("pigs.ts.csv") %>% ts()
str(pigs)
broom::glance(pigs %>% data.frame)
pigs %>% forecast::autoplot()

# forecast for the next 4 months
pigs_forecast4 <- forecast::ses(pigs, h = 4)

# tidy forecast summary
broom::tidy(pigs_forecast4 %>% summary)

# visualize output
pigs_forecast4_fitted <- fitted(pigs_forecast4)
pigs_forecast4 %>%
        autoplot() +
        autolayer(pigs_forecast4_fitted, series = "Fitted") +
        theme_classic() +
        theme(legend.position = "none") +
        xlab("Year") +
        ylab("Pigs Slaughtered")
```

### b. 95% prediction interval
```{r echo = TRUE, collapse = TRUE, fig.width = 8, fig.height = 10}
# 95% prediction interval by hand
lo.interval <- pigs_forecast4$mean[1] - 1.96 * sd(pigs_forecast4$residuals)
up.interval <- pigs_forecast4$mean[1] + 1.96 * sd(pigs_forecast4$residuals)

# R model output
lo.interval.model <- pigs_forecast4$lower[1, 2]
up.interval.model <- pigs_forecast4$upper[1, 2]

# comparison  
type <- paste0(c("calculated by hand", "calculated by hand", "run by R", "run by R"))
bound <- paste0(c("lower", "upper", "lower", "upper"))
interval <- c(lo.interval, up.interval, lo.interval.model, up.interval.model)

data.frame(type, bound, interval) %>%
        tidyr::spread(., type, interval) %>%
        dplyr::mutate(`diff (%)` = round((`calculated by hand` - `run by R`) / `calculated by hand` * 100, 2)) %>%
        kable() %>%
        kable_styling()
```

## Hyndman 7.3 {.tabset .tabset-fade .tabset-pills}

Modify your function from the previous exercise to return the sum of squared errors rather than the forecast of the next observation. Then use the `optim()` function to find the optimal values of $\alpha$ and  $\ell_0$. Do you get the same values as the `ses()` function? 

```{r echo = TRUE, collapse = TRUE}
# custom function
custom_fun <- function(fn, ts){
        a = fn[1]
        b = fn[2]
        n = length(ts)
        yhat = c(b, 0 * (2:n))
        for (i in 1:(n-1)){
                yhat[i +1] <- a * ts[i] + (1-a) * yhat[i]
        }
        res = ts - yhat
        sse = sum(res^2) / (n-1)
        
        return(sse)
}

# output parameters
parameters <- optim(par = c(0.1, 10000), fn = custom_fun, ts = pigs)
a <- parameters$par[1]
b <- parameters$par[2]

# comparison
model_a <- pigs_forecast4$model$fit$par[1]
model_b <- pigs_forecast4$model$fit$par[2]

# comparison  
type <- paste0(c("calculated by hand", "calculated by hand", "run by R", "run by R"))
parameter <- paste0(c("alpha", "level", "alpha", "level"))
value <- c(a, b, model_a, model_b)

data.frame(type, parameter, value) %>%
        dplyr::mutate(value = round(value, 3)) %>%
        tidyr::spread(., type, value) %>%
        kable() %>%
        kable_styling()
```

The result is almost identical. 
