---
title: 'XYZ'
author: 'Jimmy Ng'
output:
        pdf_document: default
        html_document:
                css: ./lab.css
                highlight: pygments
                theme: cerulean
---

### Data Preparation

```{r setup, eval = T, echo = TRUE, results = 'hide', warning = FALSE, message = FALSE}
# load packages
library(plyr)
library(lubridate)
library(tidyverse)

# load data
myfile <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-10-23/movie_profit.csv"
df <- readr::read_csv(myfile)

#
df %>%
        mutate(release_date = as.Date(release_date, "%m/%d/%Y")) %>%
        arrange(release_date) %>%
        View

```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**



### Cases 

**What are the cases, and how many are there?**



### Data collection 

**Describe the method of data collection.**



### Type of study 

**What type of study is this (observational/experiment)?**



### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**



### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**



### Independent Variable

**You should have two independent variables, one quantitative and one qualitative.**



### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```

