---
title: 'What makes a Hollywood movie profitable?'
date: "October 29, 2018"
author: 'Jimmy Ng'
output:
        pdf_document: default
        html_document: 
                css: ./lab.css
                highlight: pygments
                theme: cerulean
---

### Data Preparation
```{r setup, eval = T, echo = TRUE, collapse = TRUE, warning = FALSE, message = FALSE, prompt = TRUE, tidy = TRUE, tidy.opts = list(width.cutoff = 80)}
# load packages
library(plyr)
library(lubridate)
library(tidyverse)

# load file from github
# source from tidytuesday - they have a new dataset coming out every Tuesday
github <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data"
date <- "2018-10-23"
csv <- "movie_profit.csv"
myfile <- paste(github, date, csv, sep = "/")
df <- readr::read_csv(myfile)

###################
### have a look ###
###################
# head, dim, str
head(df)
dim(df)  # [1] 3401    9 
str(df)

############
# clean-up #
############
# check duplication: is there any movie that's duplicated in this data set?
plyr::count(df, "movie") %>%
        filter(freq >1)

# what is this movie?         
df %>%
        filter(movie == "Tau ming chong") %>%
        print
# id == 2974, 2975

# let's remove either one of these identical rows
df <- df %>%
        filter(X1 != 2974)

# is there any movie that has 0 or even negative domestic/worldwide gross?
df %>%
        filter(domestic_gross <=0 | worldwide_gross <= 0)
# there are 66 of these in this data set (as of Oct 29) 
# some of them have not been released yet, like the Aquaman!

# let's not remove them but create a flag for each of these variables
df$domestic_flag <- ifelse(df$domestic_gross <=0, 0, 1); sum(df$domestic_flag)
df$worldwide_flag <- ifelse(df$worldwide_gross <=0, 0, 1); sum(df$worldwide_flag)

# let's rename the X1 column and rename it as a movie id column
names(df)[1] <- "movie_id"

# change "date": turn the release_date column as date data type; add release day, month, year
df <- df %>%
        mutate(release_date = lubridate::mdy(release_date),
               release_day = lubridate::wday(release_date, 
                                             week_start = getOption("lubridate.week.start", 1)),
               release_month = lubridate::month(release_date),
               release_year = lubridate::year(release_date))

# rescale the production_budget, domestic_gross & worldwide_gross (by dividing 1 million)
# so that they are easier to read and visualize
df <- df %>%
        mutate(production_budget = production_budget / 1000000,
               domestic_gross = domestic_gross / 1000000,
               worldwide_gross = worldwide_gross / 1000000)
               
# change mpaa_rating & genre data type from character to factor
df <- df %>%
        mutate(mpaa_rating = factor(mpaa_rating, 
                                    levels = c("G", "PG", "PG-13", "R")),
               genre = as.factor(genre))

# complete.cases - remove all NA's
dfComplete <- df[complete.cases(df), ]

# let's look at the clean data set
head(dfComplete)
dim(dfComplete)  # [1] 3230   14
str(dfComplete)
```


### Research question 
This is a movie data set that provides various categorical and numerical variables about any major Hollywood released movie since 1936. Our main goal is to answer, "what makes a Hollywood movie profitable?" For example, do genre, mpaa rating, production budget and/or release month(s) contribute to a successful/useful multiple-regression formula in terms of predicting a movie worldwide revenue gross? Box office revenue and production budget are reported in USD and scaled to current monetary value.


### Cases 
After removing a single duplicated case, we are left with 3400 cases from the original data set; however, we need to filter out movies that are not yet released as of this moment (such as the Aquaman won't be released until December 14, 2018). In addition, we need to remove cases that have missing values. Finally, we are left with 3202 cases. 


### Data collection 
The data set (csv) is cloned from github, and it is complied from a social science project "tidytuesday" - a weekly social data project in R. As the name described, "tidytuesday" would post a data set on github every Tuesday. This movie data set is from Oct 23, 2018 and the original data is come from numbers.com - a movie industry data website that tracks box office revenue in a systematic, algorithmic way.


### Type of study 
This is an observational study with no interference of the box office. 


### Data Source 
Please see the links,

https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-23
https://thomasmock.netlify.com/post/tidytuesday-a-weekly-social-data-project-in-r/
https://www.the-numbers.com/research-analysis


### Dependent Variable
The "worldwide_gross" would be the dependent variable. It is quantitative. However, in the final project, we will likely transform and come up with a different dependent variable, e.g. "worldwide_gross / production_budget" in order to better capture the ROI of a movie.


### Independent Variable
We have the following from the data set, such as

 - "production_budget" (quantitative)
 - "mpaa_rating" (qualitative)
 - "genre" (qualitative)
 - "release_month" (qualitative - we should treat it as a seasonal factor)


### Relevant summary statistics 
Below is some summary statistics and visualization for some of the variables discussed above.


```{r relevant summary, eval = T, echo = TRUE, collapse = TRUE, warning = FALSE, message = FALSE, prompt = TRUE, tidy = TRUE, tidy.opts = list(width.cutoff = 80)}

######################################################################################
# production_budget
dfComplete %>%
        filter(worldwide_flag == 1) %>%
        select(production_budget) %>%
        summary

dfComplete %>%
        filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(y = production_budget)) +
        geom_boxplot() +
        ggtitle("Production Budget") + 
        theme_bw() + 
        labs(x = "", y = "production budget (in millions, USD)")

# worldwide_gross
dfComplete %>%
        filter(worldwide_flag == 1) %>%
        select(worldwide_gross) %>%
        summary

dfComplete %>%
        filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(y = worldwide_gross)) +
        geom_boxplot() +
        ggtitle("Worldwide Gross") + 
        theme_bw() + 
        labs(x = "", y = "worldwide gross (in millions, USD)")

######################################################################################
# table for categorical variables
# it would be interesting to create a 2 x 2 contingency table 
# or an array, in order to look at the data from a multi-dimensional perspective
# alternatively, we can do an ANOVA test on any of these with worldwide_gross
dfComplete %>%
        filter(worldwide_flag == 1) %>%
        select(mpaa_rating) %>%
        ftable

dfComplete %>%
        filter(worldwide_flag == 1) %>%
        select(genre) %>%
        ftable

dfComplete %>%
        filter(worldwide_flag == 1) %>%
        select(release_month) %>%
        ftable

######################################################################################
# in general, it seems reasonable to believe that there's a positive correlation 
# between production budget and worldwide gross

# PG, PG-13 generate the most revenue
# Action, Adventure are the most profitable movie genre 
# summer (May, June & July) is always the best time to roll out blockbuster movies!

# production_budget x worldwide_gross by mpaa_rating
dfComplete %>%
        filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = production_budget, y = worldwide_gross)) + 
        geom_point() +
        scale_y_continuous(labels = scales::dollar) +
        geom_smooth(method = "lm", col = "blue") + 
        facet_wrap(~mpaa_rating) +
        ggtitle("Worldwide Gross") + 
        labs(x = "production budget (in millions, USD)", y = "worldwide gross (in millions, USD)") +
        theme_bw() 

# production_budget x worldwide_gross by genre
dfComplete %>%
        filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = production_budget, y = worldwide_gross)) + 
        geom_point() +
        scale_y_continuous(labels = scales::dollar) +
        geom_smooth(method = "lm", col = "green") + 
        facet_wrap(~genre) +
        ggtitle("Worldwide Gross") + 
        labs(x = "production budget (in millions, USD)", y = "worldwide gross (in millions, USD)") +
        theme_bw()

# production_budget x worldwide_gross by release_month
dfComplete %>%
        filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = production_budget, y = worldwide_gross)) + 
        geom_point() +
        scale_y_continuous(labels = scales::dollar) +
        geom_smooth(method = "lm", col = "red") + 
        facet_wrap(~release_month) +
        ggtitle("Worldwide Gross") + 
        labs(x = "production budget (in millions, USD)", y = "worldwide gross (in millions, USD)") +
        theme_bw() 
```



















