---
title: 'What makes a Hollywood movie profitable?'
date: "November 14, 2018"
author: 'Jimmy Ng'
output:
        html_document: 
        pdf_document: default
        css: ./lab.css
highlight: pygments
theme: cerulean
---
        

### Data Preparation        
```{r Data Preparation, eval = T, echo = TRUE, collapse = TRUE, warning = FALSE, message = FALSE, prompt = TRUE, tidy = TRUE, warnings = F, tidy.opts = list(width.cutoff = 80)}
# load packages
library(plyr)
library(lubridate)
library(tidyverse)

# load file from github
# source from tidytuesday - they have a new dataset coming out every Tuesday
github <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data"
date <- "2018-10-23"
csv <- "movie_profit.csv"
myfile <- paste(github, date, csv, sep = "/")
df <- readr::read_csv(myfile)

###################
### have a look ###
###################
head(df)
dim(df)  # [1] 3401    9 
str(df)

############
# clean-up #
############
# check duplication: is there any movie that's duplicated in this data set?
plyr::count(df, "movie") %>%
        dplyr::filter(freq >1)

# what is this movie?         
df %>%
        dplyr::filter(movie == "Tau ming chong") %>%
        print
# id == 2974, 2975

# let's remove either one of these identical rows
df <- df %>%
        dplyr::filter(X1 != 2974)

# is there any movie that has 0 or even negative domestic/worldwide gross?
df %>%
        dplyr::filter(domestic_gross <=0 | worldwide_gross <= 0)
# there are 66 of these in this data set (as of Nov 7) 
# some of them have not been released yet, like the Aquaman!

# let's not remove them but create a flag for each of these variables
df$domestic_flag <- ifelse(df$domestic_gross <=0, 0, 1); sum(df$domestic_flag)
df$worldwide_flag <- ifelse(df$worldwide_gross <=0, 0, 1); sum(df$worldwide_flag)

# let's rename the X1 column and rename it as a movie id column
names(df)[1] <- "movie_id"

# change "date": turn the release_date column as date data type; add release day, month, year
df <- df %>%
        mutate(release_date = lubridate::mdy(release_date),
               release_day = lubridate::wday(release_date, 
                                             week_start = getOption("lubridate.week.start", 1)),
               release_month = lubridate::month(release_date),
               release_year = lubridate::year(release_date))

# rescale the production_budget, domestic_gross & worldwide_gross (by dividing 1 million)
# so that they are easier to read and/or visualize
df <- df %>%
        mutate(production_budget = production_budget / 1000000,
               domestic_gross = domestic_gross / 1000000,
               worldwide_gross = worldwide_gross / 1000000)

# change mpaa_rating & genre data type from character to factor
df <- df %>%
        mutate(mpaa_rating = factor(mpaa_rating, 
                                    levels = c("G", "PG", "PG-13", "R")),
               genre = as.factor(genre))

# complete.cases - remove all NA's
dfComplete <- df[complete.cases(df), ]

# let's look at the clean data set
head(dfComplete)
dim(dfComplete)  # [1] 3230   14
str(dfComplete)
```


### Part 1 - Introduction
The Hollywood movie industry is a multi-billion dollar industry that is so popular and influential worldwide. The industry makes a long-lasting contribution to our pop culture. A successful movie possesses the ability to capture and reshape the audiences' imagination and identities for years. It can effectively carry its messages to its audiences across time and space without restricting to any language or cultural relevancy. It's such an extraordinary industry; however, like all other industries, it is also driven by monetary gains and financial risks. 

Essentially, what makes a movie profitable? Set aside creativity and aesthetic quality, can we predict a movie's box office success purely based on quantitative variables? If so, such equation can have significant impact to transform one of the most lucrative businesses in the world.


### Part 2 - Data
Data collection: the data set (csv) is cloned from github, and it is complied from a social science project "tidytuesday" - a weekly social data project in R. As the name described, "tidytuesday" would post a data set on github every Tuesday. This movie data set is from Oct 23, 2018 and the original data is come from numbers.com - a movie industry data website that tracks box office revenue in a systematic, algorithmic way. This movie data set provides various categorical and numerical variables about any major Hollywood released movie since 1936 to present. Box office revenue and production budget are reported in USD and scaled to current monetary value.

Sources of data: 
https://github.com/rfordatascience/tidytuesday/tree/master/data/2018-10-23
https://thomasmock.netlify.com/post/tidytuesday-a-weekly-social-data-project-in-r/
https://www.the-numbers.com/research-analysis

Cases: each case in this data set represents a movie. After removing one duplicated entry, we are left with 3400 cases from the original data set; however, we need to filter out movies that are not yet released as of this moment (such as the Aquaman won't be released until December 14, 2018 in the United States). In addition, we need to remove cases that have missing values. Finally, we are left with 3230 complete cases (movies) and we are going to split it into "train" and "test" data sets.

Variables: primarily, we want to predict worldwide_gross from production_budget (numeric), mpaa_rating (factor), genre (factor), release_month (we should treat it as factor), and release_year (will group release years and transform the variable as factor).
```{r variables, eval = T, echo = T, collapse = T}
library(plyr)
variable <- lapply(dfComplete, class) %>%
        plyr::ldply(., data.frame)
variable
```

Type of study: this is an observational study with no interference of the box office. 

Scope of inference: generalizability / causality - the data is collected through historical record and this is not an experimental design. We can establish correlation, but it is not suitable to justify a causal relation in this observational study (as we do not control the production budget, genre or release month of these movies). However, the result can generate important insights for the industry. For instance, our regression model can explain how much variance that certain variables are held accountable for the box office, while holding everything else constant. We can distill from the data and figure out whether there's anything of significant interest that can impact the box office. 
    

### Part 3 - Exploratory data analysis
```{r EDA, eval = T, echo = T, collapse = T, warnings = F}
################################################
# we are interested in the following variables, 

# dependent/predicted variable: worldwide_gross
# independent/predicting variables: production_budget, mpaa_rating, genre, release_month, release_year

###################################################
# let's first look at profit and production_budget
# profit
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        dplyr::select(worldwide_gross) %>%
        summary

# production_budget
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        dplyr::select(production_budget) %>%
        summary

#################################################################################
# there are many outliners in both profit and production_budget (not surprising)
# let's do "log" transformation for both variables
dfComplete <- dfComplete %>%
        mutate(wg.log = log(worldwide_gross),
               pb.log = log(production_budget))

###########################################
# take a look of them after transformation
library(gridExtra)

## worldwide_gross ##
box.wg <- dfComplete %>% 
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(y = wg.log)) +
        geom_boxplot() +
        ggtitle("log(Worldwide Gross)") + 
        theme_bw() + 
        labs(x = "", y = "")

hist.wg <- dfComplete %>% 
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = wg.log, stat(density))) +
        geom_histogram() +
        ggtitle("log(Worldwide Gross)") + 
        theme_bw() + 
        labs(x = "", y = "")

## production_budget ##
box.pb <- dfComplete %>% 
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(y = pb.log)) +
        geom_boxplot() +
        ggtitle("log(Production Budget)") + 
        theme_bw() + 
        labs(x = "", y = "")

hist.pb <- dfComplete %>% 
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = pb.log, stat(density))) +
        geom_histogram() +
        ggtitle("log(Production Budget)") + 
        theme_bw() + 
        labs(x = "", y = "")

gridExtra::grid.arrange(box.wg, 
                        hist.wg, 
                        box.pb,
                        hist.pb,
                        ncol = 2)

##########################################
# let's look at the categorical variables
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        dplyr::select(mpaa_rating, genre, release_month, release_year) %>%
        lapply(., ftable)

# we already transformed mpaa_rating and genre from character to factor
# now let's do the same for release_month and release_year, and let's call it release_decade
dfComplete <- dfComplete %>%
        mutate(release_month = as.factor(release_month),
               release_decade = 
                       if_else(release_year < 1980, "1970s or earlier", 
                               if_else(release_year >= 1980 & release_year < 1990, "1980s", 
                                       if_else(release_year >= 1990 & release_year < 2000, "1990s", 
                                               if_else(release_year >= 2000 & release_year < 2010, "2000s", 
                                                       if_else(release_year >= 2010, "2010s", "n/a"))))),
               release_decade = as.factor(release_decade))

# let's look at these categorical variables again
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        dplyr::select(mpaa_rating, genre, release_month, release_decade) %>%
        lapply(., ftable)

#################################################################
# let's visualize the data - worldwide_gross ~ production_budget
# let's focus more on wg.log (y-axis) vs pb.log (x-axis) by different categorical variables

# wg.log vs pb.log
dfComplete %>%
    dplyr::filter(worldwide_flag == 1) %>%
    ggplot(data = ., aes(x = pb.log, y = wg.log)) + 
    geom_point() +
    geom_smooth(method = "lm", col = "grey", se = T) +
    ggtitle("log(Worldwide Gross) by log(Production Budget)") + 
    labs(x = "log(Production Budget)", y = "log(Worldwide Gross)") + 
    theme_bw()

# wg.log vs pb.log by mpaa_rating
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = pb.log, y = wg.log)) + 
        # ggplot(data = ., aes(x = production_budget, y = worldwide_gross)) + 
        geom_point() +
        geom_smooth(method = "lm", col = "blue", se = T) + 
        facet_wrap(~mpaa_rating) +
        ggtitle("MPAA Rating") + 
        labs(x = "log(Production Budget)", y = "log(Worldwide Gross)") +
        theme_bw() 

# wg.log vs pb.log by genre
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = pb.log, y = wg.log)) + 
        geom_point() +
        geom_smooth(method = "lm", col = "green", se = T) + 
        facet_wrap(~genre) +
        ggtitle("Genre") + 
        labs(x = "log(Production Budget)", y = "log(Worldwide Gross)") +
        theme_bw()

# wg.log vs pb.log by release_month
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = pb.log, y = wg.log)) + 
        geom_point() +
        geom_smooth(method = "lm", col = "red", se = T) + 
        facet_wrap(~release_month) +
        ggtitle("Release Month") + 
        labs(x = "log(Production Budget)", y = "log(Worldwide Gross)") +
        theme_bw()

# wg.log vs pb.log  by release_decade
dfComplete %>%
        dplyr::filter(worldwide_flag == 1) %>%
        ggplot(data = ., aes(x = pb.log, y = wg.log)) + 
        geom_point() +
        geom_smooth(method = "lm", col = "orange", se = T) + 
        facet_wrap(~release_decade) +
        ggtitle("Release Period") + 
        labs(x = "log(Production Budget)", y = "log(Worldwide Gross)") +
        theme_bw()
```


### Part 4 - Inference
```{r Inference, eval = T, echo = T, collapse = T, warnings = F}
###############################################################
#################### regression model #########################
###############################################################

###############################
# split the data, train + test
set.seed(1234)
dfCompleteFlag <- dfComplete %>%
        dplyr::filter(worldwide_flag == 1)
partition <- sample(x = nrow(dfCompleteFlag), 
                    size = nrow(dfCompleteFlag) * 0.7)
train <- dfCompleteFlag[partition, ]  
test <- dfCompleteFlag[-partition, ]  

###################
# build full model
# build interaction between log(production_budget) with everything else
library(broom)
fullModel <- lm(wg.log ~ pb.log * (mpaa_rating + genre + release_month + release_decade),
                data = train)
broom::tidy(fullModel) %>% 
        dplyr::filter(p.value < .05) %>% 
        dplyr::arrange(p.value) 
        
# try backward stepwise regression 
# calculate AIC
library(MASS)
backwardModel <- MASS::stepAIC(fullModel, direction = "backward")
summary(backwardModel)

########################
# (1) linearity, (2) nearly normal residuals, (3) homoscedasticity

# linear scatter plot (checked already from above EDA)
# a strong positive correlation between production budget and worldwide gross
# however, there are patterns hidden underneath the two variables
# for examples,
# PG, PG-13 generate the most revenue
# Action, Adventure draw the biggest box office, but
# Horror is the most profitable movie genre 
# summer (May, June & July) is always the best time to roll out blockbuster movies!
# therefore, we are building interaction between log(production_budget) with everything else

par(mfrow = c(1, 3))
# hist of residuals 
hist(backwardModel$residuals, main = "Histogram of residuals")
                        
# qqplot
qqnorm(backwardModel$residuals, main = "QQplot")
qqline(backwardModel$residuals)

# residual plot
plot(backwardModel$residuals ~ train$pb.log, main = "Residual chart")

par(mfrow = c(1, 1))

######################
# prediction accuracy
# apply model to test data set
testModel <- predict(backwardModel, test)

# convert back to original value
testModel <- exp(testModel)
test$predicted <- testModel

# visualize the prediction
plot(test$predicted ~ test$worldwide_gross, 
     main = "Prediction Model of Box Office Revenue\n -- dashline is only a reference line for perfect correlation",
     xlab = "actual (in USD millions)",
     ylab = "predicted (in USD millions)")
abline(a = 0, b = 1, col = "red", lty = 2)                        

# calculate RMSE (Root Mean Square Error)
library(Metrics)
rmse(actual = test$worldwide_gross, predicted = test$predicted)

####################
##### comments #####
####################
# residuals heavily skewed to the left from the histogram 
# qqplot does not support normality of the residuals either
# the final plot (actual vs. predicted) shows majority of the movies are underestimated!
# the RMSE - it is interpreted as how far on an average, the residuals are from zero
# It nullifies squared effect of MSE and provides the result in original units as data
# Let's try to minimize it!



#######################################################################
#################### rebuild regression model #########################
#######################################################################

##########################
# let's rebuild the model 
# by first removing all the outliners from the lower end of the sprectum

# recall from above EDA,
gridExtra::grid.arrange(box.wg, box.pb, ncol = 2)

# dplyr::filter outliners
wg.log.out <- boxplot(dfCompleteFlag$wg.log)$out 
pb.log.out <- boxplot(dfCompleteFlag$pb.log)$out 

dfCompleteFlag2 <- dfCompleteFlag %>%
        dplyr::filter(!(wg.log %in% wg.log.out) & !(pb.log %in% pb.log.out))
dim(dfCompleteFlag2)

#####################################
# split the data, train + test again
set.seed(1234)

partition2 <- sample(x = nrow(dfCompleteFlag2), 
                    size = nrow(dfCompleteFlag2) * 0.7)
train2 <- dfCompleteFlag2[partition2, ]  
test2 <- dfCompleteFlag2[-partition2, ]  

#########################
# build full model again
# build interaction between log(production_budget) with everything else
library(broom)
fullModel2 <- lm(wg.log ~ pb.log * (mpaa_rating + genre + release_month + release_decade),
                data = train2)
broom::tidy(fullModel2) %>% 
        dplyr::filter(p.value < .05) %>% 
        dplyr::arrange(p.value) 

# try backward stepwise regression 
# calculate AIC
library(MASS)
backwardModel2 <- MASS::stepAIC(fullModel2, direction = "backward")
summary(backwardModel2)

########################################
# regression diagnostic - check again!!
# (1) linearity, (2) nearly normal residuals, (3) homoscedasticity

# linear scatter plot (checked already)

par(mfrow = c(1, 3))

# hist of residuals 
hist(backwardModel2$residuals, main = "Histogram of residuals (ver 2)")

# qqplot
qqnorm(backwardModel2$residuals, main = "QQplot (ver 2)")
qqline(backwardModel2$residuals)

# residual plot
plot(backwardModel2$residuals ~ train2$pb.log, main = "Residual chart (ver 2)")

par(mfrow = c(1, 1))

######################
# prediction accuracy
# apply model to test data set
testModel2 <- predict(backwardModel2, test2)

# convert back to original value
testModel2 <- exp(testModel2)
test2$predicted <- testModel2

# visualize the prediction
plot(test2$predicted ~ test2$worldwide_gross, 
     main = "Prediction Model of Box Office Revenue (ver 2)\n -- dashline is only a reference line for perfect correlation",
     xlab = "actual (in USD millions)",
     ylab = "predicted (in USD millions)")
abline(a = 0, b = 1, col = "red", lty = 2)                        

# calculate RMSE (Root Mean Square Error)
library(Metrics)
rmse(actual = test2$worldwide_gross, predicted = test2$predicted)
```


### Part 5 - Conclusion
```{r conclusion, eval = T, echo = T, collapse = T, warnings = F}
summary(backwardModel2)

broom::tidy(backwardModel2) %>% 
        dplyr::filter(p.value < .05) %>% 
        dplyr::arrange(p.value) 
```

In conclusion, there's a very strong evidence that there's a positive correlation between production budget and worldwide gross. However, we must put it into the right context. It's always about an interaction between production budget AND release month or genre or release decade. Overall, the model is capable of predicting about 50% of the variance (adjusted R-squared: 0.489) with only a handful of predictors. However, after log-transformation and removing outliners, the principle of homoscedasticity still does not seem to sustain. Although ver2 of the linear model looks much better (by comparing the three diagnostic charts with the first attempt), and the AIC as well as RMSE are reduced to smaller numbers, there are still many noises in the data set that cripple the usefulness of the model. Perhaps applying a different type of GLM methodology will be better fit for this data set. In addition, we should have included more movies from the 70s and earlier, as well as 80s and 90s. The data set does not seem to be generated by a random sampling of all historically released movies. Perhaps there's a bias in data collection, i.e. including only high-profile or more well-known (but not necessarily financially successful) movies in this data set, and that skews the result and prevents us from building a useful model that can apply to the general population. A more comprehensive data set will provide tremendous help in building a better model. For example, cast list, directors, number of theatres, movie reviews, etc. 

Nevertheless, the result is still interesting and pointing to some unexpected insights. For example, we all think that summer (especially in May and June) is the best time for box office success. However, it is true only for big budget movie. In other words, releaseing a small budget movie in the summer most likely does not guarantee anything (if not resulting in box office flop). That's most likely due to the fact that audiences are more willing to go see movies in the summer, but they only rather spend their money on big budget movies (assuming better quality and CGI effects). Timing AND buget together mean everything. In addition, a surprising insight is movie genre. The most profitable genre turns out to be "Horror". Small budget horror movie is so profitable and that's why Hollywood keeps producing them! It's much safer to bet on a small budget horror movie than a small or big budget comedy or drama.

Finally, since both production budget and worldwide gross are log-transformed, that adds difficulty in interpreting the result. We must remember to convert the number after applying the model to the original data.



