---
title: "Data 624 Project 1 S04"
author: "Jeff Littlejohn"
date: "6/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# install.packages("imputeTS")
library(imputeTS) # Imputing NA \
library(xts)  # for TS object
library(forecast) # for Forecas
library(urca)
library(MLmetrics)
```



```{r}
full_data <- readxl::read_excel('Data Set for Class.xls')
head(full_data)
# For simplicity lets convert SeriesInd to date. by setting Origin of date to 1900 Jan 1st.
print(paste("Start Of Data:",as.Date(min(full_data$SeriesInd) , origin = "1900-01-01")))
print(paste("End Of Data:",as.Date(max(full_data$SeriesInd) , origin = "1900-01-01")))
print(paste("Forecast from :",as.Date(max(full_data$SeriesInd)+1 , origin = "1900-01-01")," TO ",as.Date(max(full_data$SeriesInd)+140 , origin = "1900-01-01")))
full_data$SeriesInd <- as.Date(full_data$SeriesInd, origin = "1900-01-01")
head(full_data)
```
```{r}
# Get the full date range for the data 
allDates <- seq.Date(
  min(full_data$SeriesInd),
  max(full_data$SeriesInd),
  "day")
# creating NA for missing data entry in full data.
full_data_na <- left_join(as.data.frame(allDates),full_data,by=c("allDates"= "SeriesInd"))
# subset(full_data[,c(1,2,3)], group == "S01")%>% .[,c(1,3)]  %>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
# 
# 
# subset(full_data[,c(1,2,3)], group == "S02")%>% .[,c(1,3)]  %>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
# 
# full_data_na %>% filter(.,group=="S01") %>% .[,c(1,3)] %>% plot()
# 
#   
# #+geom_col(position="dodge", alpha=0.5) 
# 
# 
# dt_s06_v1 %>% ggplot(aes(x=allDates, y= Var01))+  geom_line()
# 
# 
# 
# full_data_na[,c(1,3)]plot()
```

+  Here we breaking data by each Group and variable
- Keep Only Date and Variables value
- Merge with Full Range of Dates to see if we have any NA in the data 
```{r}
# Creating a subset of date for each Group and Varible. 
dt_s01_v1 = subset(full_data[,c(1,2,3)], group == "S01")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s01_v2 = subset(full_data[,c(1,2,4)], group == "S01")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s01_v3 = subset(full_data[,c(1,2,5)], group == "S01")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s01_v5 = subset(full_data[,c(1,2,6)], group == "S01")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s01_v7 = subset(full_data[,c(1,2,7)], group == "S01")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s02_v1 = subset(full_data[,c(1,2,3)], group == "S02")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s02_v2= subset(full_data[,c(1,2,4)], group == "S02")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s02_v3= subset(full_data[,c(1,2,5)], group == "S02")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s02_v5= subset(full_data[,c(1,2,6)], group == "S02")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s02_v7= subset(full_data[,c(1,2,7)], group == "S02")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s03_v1 = subset(full_data[,c(1,2,3)], group == "S03")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s03_v2 = subset(full_data[,c(1,2,4)], group == "S03")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s03_v3 = subset(full_data[,c(1,2,5)], group == "S03")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s03_v5 = subset(full_data[,c(1,2,6)], group == "S03")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s03_v7 = subset(full_data[,c(1,2,7)], group == "S03")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s04_v1 = subset(full_data[,c(1,2,3)], group == "S04")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s04_v2= subset(full_data[,c(1,2,4)], group == "S04")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s04_v3= subset(full_data[,c(1,2,5)], group == "S04")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s04_v5= subset(full_data[,c(1,2,6)], group == "S04")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s04_v7= subset(full_data[,c(1,2,7)], group == "S04")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s05_v1 = subset(full_data[,c(1,2,3)], group == "S04")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s05_v2 = subset(full_data[,c(1,2,4)], group == "S05")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s05_v3 = subset(full_data[,c(1,2,5)], group == "S05")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s05_v5 = subset(full_data[,c(1,2,6)], group == "S05")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s05_v7 = subset(full_data[,c(1,2,7)], group == "S05")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s06_v1 = subset(full_data[,c(1,2,3)], group == "S06")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s06_v2 = subset(full_data[,c(1,2,4)], group == "S06")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s06_v3 = subset(full_data[,c(1,2,5)], group == "S06")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s06_v5 = subset(full_data[,c(1,2,6)], group == "S06")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
dt_s06_v7 = subset(full_data[,c(1,2,7)], group == "S06")%>% .[,c(1,3)]%>% left_join(as.data.frame(allDates),.,by=c("allDates"= "SeriesInd"))
```

## S04 - Forecast  Var01, Var02

We usually see consecutive sequences of five SeriesInd entries followed by a two-integer (-days, based on our assumption) break, which might be indicative of weekdays. There are some additional breaks in the data beyond the five days on, two days off pattern. 

Note that if it were confirmed that the SeriesInd values represented weekdays, it might be prudent to remove the weekend days from the sequence for forecasting purposes.

## Exploratory Data Analysis

```{r}
summary(dt_s04_v1)
summary(dt_s04_v2)
```

We see NA's as expected. The maximum value of Var02 looks like it could be a substantial outlier.

Let's start with S04 Var01.

```{r}
plot(dt_s04_v1,main="S04 Var01")
```

There might be a seasonal drop near the end of each year, but it's not definitive. Business cycles could be present.


```{r}
plot(dt_s04_v2,main="S04 Var02")
```

There is no obvious trend or seasonality, but there are some outlier values - one clear outlier value is apparent in late 2015. It appears relatively stationary.

Because we haven't confirmed the weekend hypothesis, we will go ahead and fill the missing SeriesInd values for both Var01 and Var02 with "N/A" entries for now. We won't remove the presumed SeriesInd weekend values from the dataset.

Let's review some different methods of handling NA values in the data.

```{r}
# View of some differnet imputation methods and their values.  
cbind("Value"=dt_s01_v1$Var01,
      "na.interp"=na.interp(dt_s01_v1$Var01),
      "na.approx"=na_seadec(dt_s01_v1$Var01),
      "na.kalman"=na.kalman(dt_s01_v1$Var01),
      "na.interpolation"=na_interpolation(dt_s01_v1$Var01)
      
      ) %>% 
  head(.,20)
```

We will use the approx function.

```{r}
dt_s04_v1_xts <- xts(c(dt_s04_v1$Var01),  order.by=as.Date(dt_s04_v1$allDates))%>% na.approx()
dt_s04_v2_xts <- xts(c(dt_s04_v2$Var02),  order.by=as.Date(dt_s04_v2$allDates))%>% na.approx()
dt_s04_v1_stl <- c(dt_s04_v1$Var01,  order.by=as.Date(dt_s04_v1$allDates))%>% na.approx()
dt_s04_v2_stl <- c(dt_s04_v2$Var02,  order.by=as.Date(dt_s04_v2$allDates))%>% na.approx()
```

We appear to have two disparate datasets, so we will address them separately. Let's find the right model to forecast S04 Var01.

```{r}
ggtsdisplay(dt_s04_v1_xts,main="S04 Var01",xlab="Date",ylab="Value")
```

```{r}
dt_s04_v1_xts %>% ur.kpss() %>% summary()
```

We see clear autocorrelation between variables in the ACF. Let's take the first-order difference to see if our data can be made stationary for modeling.

```{r}
ggtsdisplay(diff(dt_s04_v1_xts),main="S04 Diff Var01",xlab="Date",ylab="Diff Value")
```

```{r}
dt_s04_v1_xts %>% diff() %>% ur.kpss() %>% summary()
```

This test statistic looks better, so we conclude that the differences data are stationary. We verify with ndiffs.

```{r}
ndiffs(dt_s04_v1_xts)
```

The S04 Var01 plots seem to indicate that ARIMA would be a good fit for this model.

```{r}
aa_fit_dt_s04_v1 <- auto.arima(coredata(dt_s04_v1_xts,stepwise=FALSE,approximation = FALSE))
summary(aa_fit_dt_s04_v1)
```




```{r}
coef(aa_fit_dt_s04_v1)
```

Auto arima did not give us drift.

```{r}
checkresiduals(aa_fit_dt_s04_v1)
```

The residuals are independent, which is what we want.

```{r}
forecast(aa_fit_dt_s04_v1,h=140) %>% autoplot()
```

```{r}
Fore_dt_s04_v1 <- forecast(aa_fit_dt_s04_v1,h=140)

autoplot(aa_fit_dt_s04_v1)
```

```{r}
Fore_dt_s04_v1
```


Now, let's find the right model to forecast S04 Var02.

```{r}
ggtsdisplay(dt_s04_v2_xts,main="S04 Var02",xlab="Date",ylab="Value")
```

```{r}
dt_s04_v2_xts %>% ur.kpss() %>% summary()
```

We see clear autocorrelation between variables in the ACF. Let's take the first-order difference to see if our data can be made stationary for modeling.

```{r}
ggtsdisplay(diff(dt_s04_v2_xts),main="S04 Diff Var02",xlab="Date",ylab="Diff Value")
```

```{r}
dt_s04_v2_xts %>% diff() %>% ur.kpss() %>% summary()
```

This test statistic looks better, so we conclude that the differences data are stationary. We verify with ndiffs.

```{r}
ndiffs(dt_s04_v2_xts)
```

The S04 Var02 plots seem to indicate that ARIMA would be a good fit for this model.

```{r}
aa_fit_dt_s04_v2 <- auto.arima(log(dt_s04_v2_xts))
summary(aa_fit_dt_s04_v2)
```

```{r}
coef(aa_fit_dt_s04_v2)
```

Auto arima did not give us drift.

```{r}
checkresiduals(aa_fit_dt_s04_v2)
```

The residuals are independent, which is what we want.

```{r}
forecast(aa_fit_dt_s04_v2,h=140) %>% autoplot()
```

```{r}
Fore_dt_s04_v2 <- forecast(aa_fit_dt_s04_v2,h=140)

autoplot(aa_fit_dt_s04_v2)
```

```{r}
fc <- forecast(aa_fit_dt_s04_v2,h=140)
fc$mean<-exp(fc$mean)
fc$upper<-exp(fc$upper)
fc$lower<-exp(fc$lower)
fc$x<-exp(fc$x)
```

```{r}
fc
```


```{r,warning=FALSE}
print(paste0("MAPE for S04 Var01 is ", MLmetrics::MAPE(Fore_dt_s04_v1$fitted,dt_s04_v1_xts)))

print(paste0("MAPE for S04 Var02 is ", MLmetrics::MAPE(exp(Fore_dt_s04_v2$fitted),dt_s04_v2_xts)))

```

```{r}
write.csv(fc,"s04v02.csv")
```

