---
title: "Multiple linear regression"
author: "Jimmy Ng"
date: "November 2, 2018"
output:
  word_document: default
  pdf_document: default
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
---

## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate 
the course and the instructor anonymously. However, the use of these student 
evaluations as an indicator of course quality and teaching effectiveness is 
often criticized because these measures may reflect the influence of 
non-teaching related characteristics, such as the physical appearance of the 
instructor. The article titled, "Beauty in the classroom: instructors' 
pulchritude and putative pedagogical productivity" (Hamermesh and Parker, 2005) 
found that instructors who are viewed to be better looking receive higher 
instructional ratings. (Daniel S. Hamermesh, Amy Parker, Beauty in the 
classroom: instructors pulchritude and  putative pedagogical productivity, 
*Economics of Education Review*, Volume 24, Issue 4, August 2005, Pages 369-376, 
ISSN 0272-7757, 10.1016/j.econedurev.2004.07.013. [http://www.sciencedirect.com/science/article/pii/S0272775704001165](http://www.sciencedirect.com/science/article/pii/S0272775704001165).)

In this lab we will analyze the data from this study in order to learn what goes 
into a positive professor evaluation.

## The data

The data were gathered from end of semester student evaluations for a large 
sample of professors from the University of Texas at Austin. In addition, six 
students rated the professors' physical appearance. (This is aslightly modified 
version of the original data set that was released as part of the replication 
data for *Data Analysis Using Regression and Multilevel/Hierarchical Models* 
(Gelman and Hill, 2007).) The result is a data frame where each row contains a 
different course and columns represent variables about the courses and professors.

```{r load-data, eval=TRUE}
load("more/evals.RData")
```

variable         | description
---------------- | -----------
`score`          | average professor evaluation score: (1) very unsatisfactory - (5) excellent.
`rank`           | rank of professor: teaching, tenure track, tenured.
`ethnicity`      | ethnicity of professor: not minority, minority.
`gender`         | gender of professor: female, male.
`language`       | language of school where professor received education: english or non-english.
`age`            | age of professor.
`cls_perc_eval`  | percent of students in class who completed evaluation.
`cls_did_eval`   | number of students in class who completed evaluation.
`cls_students`   | total number of students in class.
`cls_level`      | class level: lower, upper.
`cls_profs`      | number of professors teaching sections in course in sample: single, multiple.
`cls_credits`    | number of credits of class: one credit (lab, PE, etc.), multi credit.
`bty_f1lower`    | beauty rating of professor from lower level female: (1) lowest - (10) highest.
`bty_f1upper`    | beauty rating of professor from upper level female: (1) lowest - (10) highest.
`bty_f2upper`    | beauty rating of professor from second upper level female: (1) lowest - (10) highest.
`bty_m1lower`    | beauty rating of professor from lower level male: (1) lowest - (10) highest.
`bty_m1upper`    | beauty rating of professor from upper level male: (1) lowest - (10) highest.
`bty_m2upper`    | beauty rating of professor from second upper level male: (1) lowest - (10) highest.
`bty_avg`        | average beauty rating of professor.
`pic_outfit`     | outfit of professor in picture: not formal, formal.
`pic_color`      | color of professor's picture: color, black & white.

## Exploring the data

1.  Is this an observational study or an experiment? The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question.

# JN: strictly speaking, this is an observational study because we do not control about assigning students to different professors/classes. What if all English professors are considered to be significantly more attractive than professors teaching mathematics? Neither do we control the "attractiveness" of professors in this student. We cannot do randomized control trial. However, we can still answer the question by carrying out a correlation or quasi-experimental study.

2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not?
```{r, eval = T, echo = T, collapse = T}
par(mfrow = c(1, 2))
hist(evals$score, main = "histogram of score")
boxplot(evals$score, main = "boxplot of score")
```

# JN: Yes that's heavily left skewed. That means majority of students have positive rating of their professors, with only few exceptions that rate their professors negatively (or below average). That's what I expected - majority of students have positive learning experience. 

3.  Excluding `score`, select two other variables and describe their relationship 
    using an appropriate visualization (scatterplot, side-by-side boxplots, or 
    mosaic plot).
```{r, eval = T, echo = T, collapse = T}
with(evals, 
     plot(bty_avg ~ age))
with(evals,
     cor.test(bty_avg, age))
```

# JN: there's a very weak negative correlation between a professor's age and his/her average beauty rating. 

## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers
are evaluated more favorably. Let's create a scatterplot to see if this appears 
to be the case:

```{r scatter-score-bty_avg, eval = FALSE}
plot(evals$score ~ evals$bty_avg)
```


Before we draw conclusions about the trend, compare the number of observations 
in the data frame with the approximate number of points on the scatterplot. 
Is anything awry?

4.  Replot the scatterplot, but this time use the function `jitter()` on the
    $y$- or the $x$-coordinate. (Use `?jitter` to learn more.) What was
    misleading about the initial scatterplot?
```{r, eval = FALSE}
plot(evals$score ~ jitter(evals$bty_avg))
```

# JN: what's misleading about the scatterplot is that, although the two variables look like "continuous", they are actually not. If we run library(tidyverse), and then "unique(evals$bty_avg) %>% length" and "unique(evals$bty_avg) %>% length", we will see that we only have 27 uniques for score and 35 for bty_avg. In other words, there are "spaces" or "gaps" between dots and as a result, the scatterplot depicts the relationship between the two variables visually less convincing. 

5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating and add the line to your plot
    using `abline(m_bty)`. Write out the equation for the linear model and
    interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?
```{r eval = T, echo = T, collapse = T}
m_bty <- lm(score ~ bty_avg, data = evals)
plot(jitter(evals$score) ~ jitter(evals$bty_avg))
abline(m_bty, lty = 2, col = "red")
summary(m_bty)
```

# JN: the equation is y = 0.06664 * x + 3.88034 // the slope (0.06664) is positive, meaning every single increase in bty_avg would bring an increase of score by 0.06664. // Yes, the bty_avg is a significant predictor with p-value far below 0.05, i.e. 5.08e-05. However, it may not be practically significant or useful because the estimate is so small (0.06664).

6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).
```{r eval = T, echo = T}
plot(m_bty$residuals ~ evals$bty_avg)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
```

# JN: yes the condition is met. The above residual plot shows a random pattern - the randomness indicating the satisfaction of homoscedasticity.

## Multiple linear regression

The data set contains several variables on the beauty score of the professor: 
individual ratings from each of the six students who were asked to score the 
physical appearance of the professors and the average of these six scores. Let's 
take a look at the relationship between one of these scores and the average 
beauty score.

```{r bty-rel, eval = FALSE}
plot(evals$bty_avg ~ evals$bty_f1lower)
cor(evals$bty_avg, evals$bty_f1lower)
```

As expected the relationship is quite strong - after all, the average score is 
calculated using the individual scores. We can actually take a look at the 
relationships between all beauty variables (columns 13 through 19) using the 
following command:

```{r bty-rels, eval = FALSE}
plot(evals[,13:19])
```

These variables are collinear (correlated), and adding more than one of these 
variables to the model would not add much value to the model. In this 
application and with these highly-correlated predictors, it is reasonable to use
the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score 
after we've accounted for the gender of the professor, we can add the gender 
term into the model.

```{r scatter-score-bty_avg_gender, eval = T, echo = T}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.
```{r, eval = T, echo = T}
library(car)
par(mfrow = c(1, 2))
hist(m_bty_gen$residuals, main = "histogram of the residuals")
car::qqPlot(m_bty_gen, main = "qqplot of the residuals")
```

# JN: the residuals are nearly normally distributed (with left skewed).  

8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?

# JN: Yes, it still is a significant predictor with p-value 6.48e-06. Adding "gender" to the model also changed the parameter estimate of "bty_avg" from 0.06664 to 0.07416.

Note that the estimate for `gender` is now called `gendermale`. You'll see this 
name change whenever you introduce a categorical variable. The reason is that R 
recodes `gender` from having the values of `female` and `male` to being an 
indicator variable called `gendermale` that takes a value of $0$ for females and
a value of $1$ for males. (Such variables are often referred to as "dummy" 
variables.)

As a result, for females, the parameter estimate is multiplied by zero, leaving 
the intercept and slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

We can plot this line and the line corresponding to males with the following 
custom function.

```{r twoLines, eval = FALSE}
multiLines(m_bty_gen)
```

9.  What is the equation of the line corresponding to males? (*Hint:* For
    males, the parameter estimate is multiplied by 1.) For two professors
    who received the same beauty rating, which gender tends to have the
    higher course evaluation score?
    
# JN: y = 3.74734 + x1 * (0.07416) + 1 * (0.17239). "Male" would tend to have higher score. If that's "female", the estimate (0.17239) would be multiplied by 0, and that means it will not be added or contributed to the increase in score.

The decision to call the indicator variable `gendermale` instead of`genderfemale`
has no deeper meaning. R simply codes the category that comes first 
alphabetically as a $0$. (You can change the reference level of a categorical 
variable, which is the level that is coded as a 0, using the`relevel` function. 
Use `?relevel` to learn more.)

10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.
```{r eval = T, echo = T, collapse = T}
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)
```

# JN: the categorical variable would be "dummified". If it has 3 levels, then you will see it appear twice (k - 1) as shown above. 

The interpretation of the coefficients in multiple regression is slightly 
different from that of simple regression. The estimate for `bty_avg` reflects
how much higher a group of professors is expected to score if they have a beauty
rating that is one point higher *while holding all other variables constant*. In
this case, that translates into considering only professors of the same rank 
with `bty_avg` scores that are one point apart.

## The search for the best model

We will start with a full model that predicts professor score based on rank, 
ethnicity, gender, language of the university where they got their degree, age, 
proportion of students that filled out evaluations, class size, course level, 
number of professors, number of credits, average beauty rating, outfit, and 
picture color.

11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score.
    
# JN: I expect that cls_profs would have the highest p-value, simply because the number of professors teaching the course should have nothing to do with the score.

Let's run the model...

```{r m_full, eval = FALSE, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.
    
# JN: Yes, my suspicion is confirmed - cls_profssingle has the highest p-value of "0.77806". That means, the variable is highly irrelevant to the regression model. 

13. Interpret the coefficient associated with the ethnicity variable.

# JN: once again, the p-value is not significant, i.e. 0.11698, meaning ethnicity of professor does not significantly impact the score. To interpret the coefficient, we can state that as when a professor is "not minority", we would add 0.1234929 to the score, holding everything else constant.

14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?
```{r eval = FALSE, tidy = FALSE}
m_parital <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_parital)
```

# JN: Yes, the coefficients and significance of the other explanatory variables change after dropping "cls_profs". 

15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.
```{r backward selection, eval = T, echo = T, collapse = T}
m_all <- lm(score ~., data = evals)

library(MASS)
MASS::stepAIC(m_all, direction = "backward")
```

# JN: as shown above, we can use the *AIC* (Akaike's Information Criteria) to determine the best model. Using direction = "backward" and starting with m_all (by including every explanatory variable), we found the best model (or stopped removing variable) with the smallest AIC returned.

And the model is described below,

lm(formula = score ~ rank + ethnicity + gender + language + age + 
    cls_perc_eval + cls_credits + bty_f1lower + bty_f1upper + 
    bty_f2upper + bty_avg + pic_outfit + pic_color, data = evals)



16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.
```{r evaluation, eval = T, echo = T, collapse = T}
m_final <- lm(formula = score ~ rank + ethnicity + gender + language + age + 
           cls_perc_eval + cls_credits + bty_f1lower + bty_f1upper + 
           bty_f2upper + bty_avg + pic_outfit + pic_color, data = evals)
summary(m_final)

par(mfrow = c(1, 3))
hist(m_final$residuals, main = "histogram of the residuals")
car::qqPlot(m_final, main = "qqplot of the residuals")
plot(m_final$residuals, main = "residuals' plot")
```

# JN: Yes, every condition is checked out!!!

17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?

# JN: it won't, consider the sample (in this case, professors) were randomly selected. There's no reason to believe that the courses that were taught by these random samples contain bias or misrepresent the general population.

18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.

# JN: Most likely, a professor is male, teaching, not minority, received education from school that was taught in English, relatively young and less likely to be rated high in bty_avg. 

19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?

# JN: I would feel uncomfortable because the adjusted R-squared is only 0.1821. That's too weak a relationship between all independent variables and the dependent variable. There is still over 80% of the variance not explained by the model. We need to collect different data or redesign our study by coming up with better hypothesis for testing.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was written by 
Mine &Ccedil;etinkaya-Rundel and Andrew Bray.
</div>
